Model Estimation

### 2018-12-05 model_estimation.Rmd
Iteratively optimize function with tweets and with no tweets

```{r tweettransformation}
parks@data %>%
  tbl_df() %>%
  select(raw = TWEET_COUNT) %>%
  filter(raw < 100) %>%
  mutate(
    `0.5` = yeo.johnson(raw, 0.5),
    `0` = yeo.johnson(raw, 0),
    log = log(raw)
  ) %>%
  gather(lambda, x, -raw) %>%
  ggplot(aes(x = raw, y = x, color = lambda)) + 
  geom_line()
```

```{r show_data, echo=FALSE}
tracts@data <- tbl_df(tracts@data) %>%
  select(
    GEOID, OBESITY, Park_Percent, Phys_Act, MENTAL, 
    Income1, Income2, Income3, Income4, Income5, Income6, 
    Income7, Income8, Income9, Income10, Pop_Density, 
    FulltimeWork, CollegeDeg, Pct0to17, Pct18to29, Pct30to64, Pct65plus, 
    Single_Percent, PctWhite, PctBlack, PctNative, PctAsian, PctPacific,
    PctOther, PctHispanic
  )

@@

```{r spatial selection}
obese_base_lm <- lm(update(base_formula, OBESITY ~ .), data = tracts@data)
obese_base_sar <- lagsarlm(update(base_formula, OBESITY ~ .), 
                           data = tracts@data, listw = W, zero.policy = TRUE)
obese_base_sem <- errorsarlm(update(base_formula, OBESITY ~ .), 
                             data = tracts@data, listw = W, zero.policy = TRUE)
obese_base_sdm <- lagsarlm(update(base_formula, OBESITY ~ .), 
                           data = tracts@data, listw = W, zero.policy = TRUE, 
                           type = "mixed")
```

@@

# distances from all tracts to all zones
distances <- gDistance(park_centroids, tract_centroids, byid = TRUE) / 5280
```

Now we want to generate an experiment where we supply different $\beta$ values
to the logsum calculation in order to maximize the log-likelihood function of the
regression model.

## Model with tweets

```{r tweets_logsum}
tweets_logsum <- function(betas, distances, parks, tracts) {

  logsum <- calculate_park_logsums(
    log(distances), 
    log(parks$Park_Acres), 
    yeo.johnson(parks$TWEET_COUNT, lambda = 0),
    betas = betas
  )

  estimation_data <- tracts@data %>% mutate(logsum = logsum)

  # sdm <- lagsarlm(
  #   update(base_formula, OBESITY ~ . + logsum),   
  #   data = estimation_data,  listw = W, zero.policy = TRUE, type = "mixed")
  sdm <- lm(update(base_formula, OBESITY ~ . + logsum), data = estimation_data)

@@

tweets_lm_val <- optim(fn = tweets_logsum, 
                       par = c(-1, 0, 0), # distance, size, tweets
                       lower = c(-Inf, 0, -Inf), # size coefficient must be positive
                       upper = c(0, Inf, Inf), # distance coefficient must be negative
                       hessian=TRUE, 
                       distances = distances, parks = parks, tracts = tracts,
                       method = "L-BFGS-B", control = list(trace = 6))
```

```{r tweets_sdm}
tweets_ls <- calculate_park_logsums(
    log(distances), 
    log(parks$Park_Acres), 
    yeo.johnson(parks$TWEET_COUNT, lambda = 0),
    betas = tweets_lm_val$par)
tracts$tweets_ls <- tweets_ls - mean(tweets_ls)
tweets_sdm <- lagsarlm(update(base_formula, OBESITY ~ . + tweets_ls),  
         data = tracts@data,  listw = W, zero.policy = TRUE, type = "mixed")
```

```{r notweets_logsum}
notweets_logsum <- function(betas, distances, parks, tracts) {

  logsum <- calculate_park_logsums(
    log(distances), 
    log(parks$Park_Acres), 
    betas = betas
  )

  estimation_data <- tracts@data %>% mutate(logsum = logsum)

  # sdm <- lagsarlm(
  #   update(base_formula, OBESITY ~ . + logsum),   
  #   data = estimation_data,  listw = W, zero.policy = TRUE, type = "mixed")
  sdm <- lm(update(base_formula, OBESITY ~ . + logsum), data = estimation_data)

notweets_lm_val <- optim(fn = notweets_logsum, 
                       par = c(-1, 0), # distance, size
                       lower = c(-100, 0), # size coefficient must be positive
                       upper = c(0, 100), # distance coefficient must be negative
                       hessian=TRUE, 
                       distances = distances, parks = parks, tracts = tracts,
                       method = "L-BFGS-B", control = list(trace = 6))
```

```{r tweets_sdm}
notweets_ls <- calculate_park_logsums( 
  log(distances),  
  log(parks$Park_Acres),  
  betas = notweets_lm_val$par)

tracts@data$notweets_ls <- notweets_ls - mean(notweets_ls)

notweets_sdm <- lagsarlm(update(base_formula, OBESITY ~ . + notweets_ls),  
         data = tracts@data,  listw = W, zero.policy = TRUE, type = "mixed")
```

```{r modelchart}
obesity_models <- list(Base = obese_base_sdm, `Access` = notweets_sdm, 
                       `Access + Tweets` = tweets_sdm)
screenreg(obesity_models, digits = 4)
```

@@

  coef <- list(
    direct = impacts_summary$direct_sum,
    indirect = impacts_summary$indirect_sum,
    total = impacts_summary$total_sum
  ) %>%
  lapply(function(s) {
    data_frame(
      var = names(s$statistics[,1]),
      impact = s$statistics[,1]
    )
  }) %>%
    bind_rows(.id = "effect")

  pval <- as_data_frame(impacts_summary$pzmat) %>%
    mutate(
     var = rownames(impacts_summary$pzmat)
    ) %>%
    gather("effect", "p-val", -var) %>%
    mutate(effect = tolower(effect))

@@

```{r physact_estimation}
# turn around the metric to make sense
tracts$PHYSACT = 100 - tracts$Phys_Act
pa_base_sdm <- lagsarlm(update(base_formula, PHYSACT ~ .),
         data = tracts@data,  listw = W, zero.policy = TRUE, type = "mixed")
pa_notweets_sdm <- lagsarlm(update(base_formula, PHYSACT ~ . + notweets_ls),  
          data = tracts@data,  listw = W, zero.policy = TRUE, type = "mixed")
pa_tweets_sdm <- lagsarlm(update(base_formula, PHYSACT ~ . + tweets_ls),  
          data = tracts@data,  listw = W, zero.policy = TRUE, type = "mixed")
```

@@

>>>>

```{r tweettransformation}
parks@data %>%
  tbl_df() %>%
  dplyr::select(raw = TWEET_COUNT) %>%
  filter(raw < 100) %>%
  mutate(
    `0.5` = yeo.johnson(raw, 0.5),
    `0` = yeo.johnson(raw, 0),
    log = log(raw)
  ) %>%
  gather(lambda, x, -raw) %>%
  ggplot(aes(x = raw, y = x, color = lambda)) + 
  geom_line()

parks@data <- parks@data %>%
  mutate(
    log_size = log(Park_Acres),
    log_tweets = yeo.johnson(TWEET_COUNT, 0)
  )

```

## Other variables

```{r show_data, echo=FALSE}
tracts@data <- tbl_df(tracts@data) %>%
  dplyr::select(
    GEOID, OBESITY, Park_Percent, Phys_Act, MENTAL, 
    Income1, Income2, Income3, Income4, Income5, Income6, 
    Income7, Income8, Income9, Income10, Pop_Density, 
    FulltimeWork, CollegeDeg, Pct0to17, Pct18to29, Pct30to64, Pct65plus, 
    Single_Percent, PctWhite, PctBlack, PctNative, PctAsian, PctPacific,
    PctOther, PctHispanic
  )  %>%
  mutate(
    log_obesity = log(OBESITY),
    log_physact = log(100 - Phys_Act)
  )

@@

```{r spatial selection}
obese_base_lm <- lm(update(base_formula, log_obesity ~ .), data = tracts@data)
obese_base_sar <- lagsarlm(update(base_formula, log_obesity ~ .), 
                           data = tracts@data, listw = W, zero.policy = TRUE)
obese_base_sem <- errorsarlm(update(base_formula, log_obesity ~ .), 
                             data = tracts@data, listw = W, zero.policy = TRUE)
obese_base_sdm <- lagsarlm(update(base_formula, log_obesity ~ .), 
                           data = tracts@data, listw = W, zero.policy = TRUE, 
                           type = "mixed")
```

@@

# distances from all tracts to all zones
distances <- log(gDistance(park_centroids, tract_centroids, byid = TRUE) / 5280)
```

Now we want to generate an experiment where we supply different $\beta$ values
to the logsum calculation in order to maximize the log-likelihood function of the
regression model. This will help us identify the proper coefficients to use for
the logsum construction.

```{r notweets_logsum}
notweets_logsum <- function(betas, distances, parks, tracts, model) {

  estimation_data <- tracts@data %>% 
    mutate(logsum = calculate_park_logsums(
      distances,  parks$log_size,
      betas = betas))

  sdm <- lagsarlm(update(model, . ~ . + logsum), data = estimation_data, 
                  listw = W, zero.policy = TRUE, type = "mixed")
  return(-as.numeric(logLik(sdm)))
}

access_sdm_val <- optim(
  fn = notweets_logsum,  
  par = c(-1, 0), # distance, size 
  lower = c(-100, 0), # size coefficient must be positive 
  upper = c(0, 100), # distance coefficient must be negative 
  hessian=TRUE, distances = distances, parks = parks, tracts = tracts, 
  model = obese_base_lm, method = "L-BFGS-B", control = list(trace = 6))
```

```{r tweets_logsum}
tweets_logsum <- function(betas, distances, parks, tracts, model) {

  estimation_data <- tracts@data %>% 
    mutate(logsum = calculate_park_logsums(
      distances,  parks$log_size,  parks$log_tweets,
      betas = betas))

  sdm <- lagsarlm(update(model, . ~ . + logsum), data = estimation_data, 
                  listw = W, zero.policy = TRUE, type = "mixed")

@@

tweets_sdm_val <- optim(
  fn = notweets_logsum,  
  par = c(-1, 1, 0),
  lower = c(-100, 0, -100), # size coefficient must be positive
  upper = c(-1e-9, 100, 100), # distance coefficient must be negative
  hessian=TRUE, distances = distances, parks = parks, tracts = tracts, 
  model = obese_base_lm, method = "L-BFGS-B", control = list(trace = 6))

```

```{r obesity_access_models}
tracts@data <- tracts@data %>%
  mutate(
    access_ls = calculate_park_logsums(distances, parks$log_size, 
                                       betas = notweets_lm_val$par),
    tweets_ls = calculate_park_logsums(distances, parks$log_size, parks$log_tweets, 
                                       betas = tweets_lm_val$par),
    # demean
    access_ls = access_ls - mean(access_ls),
    tweets_ls = tweets_ls - mean(tweets_ls)
  )

access_sdm <- lagsarlm(update(obese_base_sdm, .~ . + access_ls),   
                       data = tracts@data,  listw = W, zero.policy = TRUE, 
                       type = "mixed")
tweets_sdm <- lagsarlm(update(obese_base_sdm, .~ . + tweets_ls),   
                       data = tracts@data,  listw = W, zero.policy = TRUE, 
                       type = "mixed")
```

```{r modelchart}
obesity_models <- list(Base = obese_base_sdm, `Access` = access_sdm, 
                       `Access + Tweets` = tweets_sdm)
screenreg(obesity_models, digits = 4)
```

@@

  coef <- list(
    Direct = impacts_summary$direct_sum,
    Indirect = impacts_summary$indirect_sum,
    Total = impacts_summary$total_sum
  ) %>%
  lapply(function(s) {
    data_frame(
      var = names(s$statistics[,1]),
      impact = s$statistics[,1]
    )
  }) %>%
    bind_rows(.id = "effect")

  pval <- as_data_frame(impacts_summary$pzmat) %>%
    mutate(
     var = rownames(impacts_summary$pzmat)
    ) %>%
    gather("effect", "p-val", -var)

@@

```{r physact_estimation}
pa_base_sdm <- lagsarlm(update(base_formula, PHYSACT ~ .),
         data = tracts@data,  listw = W, zero.policy = TRUE, type = "mixed")
pa_access_sdm <- lagsarlm(update(base_formula, PHYSACT ~ . + notweets_ls),  
          data = tracts@data,  listw = W, zero.policy = TRUE, type = "mixed")
pa_tweets_sdm <- lagsarlm(update(base_formula, PHYSACT ~ . + tweets_ls),  
          data = tracts@data,  listw = W, zero.policy = TRUE, type = "mixed")
```

@@

# Results
```{r impacts_table}
list("Obesity" = obesity_impacts, "Physical activity" = physact_impacts) %>%
  bind_rows(.id = "Outcome") %>%
  filter(var %in% c("access_ls", "tweets_ls")) %>%
  dplyr::select(-var, -Base)  %>%
  gather("Variable", "Impact", -Outcome, -effect) %>%
  filter(!is.na(Impact))  %>%
  spread(effect, Impact)

```

```{r logsum_scales}
data_frame(
  Parameter = c("Distance", "Size", "Tweets"),
  Access = c(notweets_lm_val$par, NA),
  `Access + Tweets` = tweets_lm_val$par
)

```
